We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-07 paper/paper.tex
typos

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ greatly among the simulations, suggesting a
strong influence of resolution, simulated physics and criteria to select
LG pairs.
However, there are three common trends in these simulations. 
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity aas the Local Group; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW is atypical in
this respect.  
These results place the LG at the level of a $3\sigma$ outlier
in the LCDM paradigm. 
The approach presented here facilitates the comparison between
asphericity results from different numerical setups and choices to
select halo pairs in simulations, hopefully helping to clarify the
reasons behin the LG atipicallity in LCDM. 
\end{abstract}

@@

For both the M31 and MW we only use the 11 to 15 brightest satellites(using
$M_V$ magnitudes) within a distance of $300$kpc to its central galaxy.
The satellites included for the MW analysis are: 
LMC, SMC, Canis Major, Sagittarius dSph, Fornax, Leo I, Sculptor,
Leo II, Sextans I, Carina, Ursa Minor, Draco, Canes Venatici (I),
Hercules and Bootes II.
The satellites included for the MW analysis are: Triangulum, NGC205,
M32, IC10, NGC185, NGC147, Andromeda VII, Andromeda II, Andromeda
XXXII, Andromeda XXXI, Andromeda I, Andromeda VI, Andromeda XXIII, LGS 3, 
 and Andromeda III.

@@

In this paper we develop and demostrate a method to quantify the
asphericity of the satellite distribution in the Local Group.
We build up on the suggestions made in \cite{XX}.
We do not look for planes or coherent kinematical structures in the
interest of keeping the method straightforward and robust.

>>>>

\begin{abstract}
We use numerical simulations to build an explicit probability
distribution for the asphericity in the satellite distribution around
galaxies similar to the Local Group (LG) in the Lambda Cold Dark
Matter (LCDM) paradigm. 
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-Dark and ELVIS. 
Detailed results differ greatly among the simulations, suggesting a
strong influence of the typical DM halo mass in the different pair samples.
However, there are three common trends in these simulations. 
First, at most of $2\%$ the pairs are expected to have satellite
distributions with the same asphericity as the Local Group; second,
between $27\%$ to $56\%$ of the pairs have a halo with a satellite
distribution as aspherical as in M31; and third, at most $3\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The brightest satellites in M31 galaxy turn out to have a rather
typical distribution in the LCDM context, while the MW is atypical in
this respect.  
These results place the LG at the level of a $3\sigma$ outlier
in the LCDM paradigm. 
The approach presented here facilitates the comparison between
asphericity results from different numerical setups and choices to
select halo pairs in simulations, hopefully helping out to clarify the
reasons behind the LG atipicallity in LCDM. 
\end{abstract}

@@

For both the M31 and MW we only use the 11 to 15 brightest satellites(using
$M_V$ magnitudes) within a distance of $300$kpc to its central galaxy.
The satellites included for the MW analysis are: 
LMC, SMC, Canis Major, Sagittarius dSph, Fornax, Leo I, Sculptor,
Leo II, Sextans I, Carina, Ursa Minor, Draco, Canes Venatici (I),
Hercules and Bootes II.
The satellites included for the M31 analysis are: Triangulum, NGC205,
M32, IC10, NGC185, NGC147, Andromeda VII, Andromeda II, Andromeda
XXXII, Andromeda XXXI, Andromeda I, Andromeda VI, Andromeda XXIII, LGS 3, 
 and Andromeda III.

@@

In this paper we develop and demostrate a method to quantify the
asphericity of the satellite distribution in the Local Group.
We build up on the suggestions made in \cite{2017AN....338..854P}.
We do not look for planes or coherent kinematical structures in the
interest of keeping the method straightforward and robust.
